$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-06 paper/paper.tex
Add "(Sketch)" to SR proof
Use subparagraph title as the start of intro sentences

\subparagraph*{Framing}
Framing asserts that the well-typedness of a process is independent of its leftover resources.

@@

\subparagraph*{Weakening}
Weakening states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
\begin{nitheorem}[Weakening]
  \label{thm:weakening}
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, lifting every variable greater than or equal to $i$ in $P$ is well typed in
  $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
\end{nitheorem}

\subparagraph*{Strengthening}
Removing an unused variable preserves the well-typedness of a process.
Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
\begin{nitheorem}[Strengthening]
  \label{thm:strengthening} 
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Let $i$ be a variable not in $P$, such that $\Unused_i \; P$.
  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
\end{nitheorem}

@@

\subparagraph*{Swapping}
Swapping two variables preserves the well-typedness of a process.
For simplicity, we extend $\func{swap}_i$ introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
\begin{nitheorem}[Swapping]
  \label{thm:swapping}
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\allowbreak \func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}

@@

\subparagraph*{Subject Congruence}
Subject congruence states that applying structural congruence rules to a well typed process preserves its well-typedness.

@@

\subparagraph*{Substitution}
\autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.

@@

\begin{proof}[Proof (Sketch)]
  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  Full details are found in \autoref{app:subst-generalization}.
  \begin{itemize}
   \item
     Constructor $\constr{end}$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
   \item
     Constructor $\constr{chan}$: we proceed inductively.
   \item
     Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction.
 \end{itemize}
\end{proof}

@@

\subparagraph*{Subject Reduction}
Subject reduction states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).

@@

\begin{proof}
  By induction on $P \reduce{c} Q$.
  \hfill{}\\
  \begin{itemize}
    \item
    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).

>>>>

\subparagraph*{Framing}
asserts that the well-typedness of a process is independent of its leftover resources.

@@

\subparagraph*{Weakening}
states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
\begin{nitheorem}[Weakening]
  \label{thm:weakening}
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, lifting every variable greater than or equal to $i$ in $P$ is well typed in
  $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
\end{nitheorem}

\subparagraph*{Strengthening}
states that removing an unused variable preserves the well-typedness of a process.
Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
\begin{nitheorem}[Strengthening]
  \label{thm:strengthening} 
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Let $i$ be a variable not in $P$, such that $\Unused_i \; P$.
  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
\end{nitheorem}

@@

\subparagraph*{Swapping}
two variables preserves the well-typedness of a process.
For simplicity, we extend $\func{swap}_i$ introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
\begin{nitheorem}[Swapping]
  \label{thm:swapping}
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\allowbreak \func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}

@@

\subparagraph*{Subject Congruence}
states that applying structural congruence rules to a well typed process preserves its well-typedness.

@@

\subparagraph*{Substitution}
\todo{intro}
\autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.

@@

\begin{proof}[Proof (Sketch)]
  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  Full details are found in \autoref{app:subst-generalization}.
  \begin{itemize}
   \item
     Constructor $\constr{end}$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
   \item
     Constructor $\constr{chan}$: we proceed inductively.
   \item
     Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qedhere
 \end{itemize}
\end{proof}

@@

\subparagraph*{Subject Reduction}
states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).

@@

\begin{proof}[Proof (Sketch)]
  By induction on $P \reduce{c} Q$.
  \hfill{}\\
  \begin{itemize}
    \item
    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
